#Peer Assessment 1
```{r echo=FALSE}
# Load libraries
library(plyr)
```

###Loading and preprocessing the data

* First we need to load the activity data that we will be analyzing
```{r loadData}
# Read data using read.csv()
data <- read.csv("activity.csv", na.strings=c("NA"))
```

###What is the mean total number of steps taken per day?

* Calculate the total number of steps taken per day:
```{r}
# Remove NA values
data <- data[is.na(data$steps)==0,]

# Summarize steps by day using plyr library
dailySteps <- ddply(data, .(date), summarize, steps=sum(steps))
```

* Histogram of steps taken per day:

```{r echo=FALSE}
hist(dailySteps$steps, xlab="Steps", main="Histogram of Steps Taken Per Day")
```

* Calculate the mean and median of the total steps taken per day

```{r}
# Calculate mean of steps taken per day
meanDailySteps <- format(round(mean(dailySteps$steps), digits=2), nsmall=2)

# Calculate median of steps taken per day
medianDailySteps <- as.integer(median(dailySteps$steps))
```

* The mean for daily steps = `r meanDailySteps`
* The median for daily steps = `r medianDailySteps`

###What is the average daily activity pattern?

* Calculate the average number of steps taken per 5 minute interval
```{r}
# Summarize steps per interval
intervalSteps <- ddply(data, .(interval), summarize, steps=mean(steps))
```

* Plot of average number of steps taken per interval
```{r echo=FALSE}
plot(intervalSteps$interval, intervalSteps$steps, type="l", xlab="Interval", ylab="Steps")
```

```{r echo=FALSE}
maxInterval <- head(intervalSteps[order(-intervalSteps$steps),], 1)
```

* The 5 minute interval with the highest average number of steps is **`r maxInterval$interval`** with an average of **`r maxInterval$steps` steps**.

###Imputting missing values

* Let's load up all the data again including the missing values:
```{r ref.label='loadData'}
```

* Let's find all of the missing values
```{r}
# Find the incomplete cases and count
nrow(data[complete.cases(data)==0,])
```

* Now let's get the mean per interval to replace the missing values
```{r}
# Calculate the mean per interval
meanStepsPerInterval <- ddply(data[complete.cases(data),], .(interval), summarize, meanSteps=mean(steps))

# Merge data frames
data <- merge(data, meanStepsPerInterval, by.x="interval", by.y="interval")

# Set steps to meanSteps where steps is NA
data[is.na(data$steps),]$steps <- data[is.na(data$steps),]$meanSteps

# Summarize steps by day using plyr library
newDailySteps <- ddply(data, .(date), summarize, steps=sum(steps))
```

* Here's a histogram of the data set with the missing values replaced:
```{r echo=FALSE}
hist(newDailySteps$steps, xlab="Steps", main="Histogram of Steps Taken Per Day")
```

* Calculate the mean and median of the total steps taken per day for the new data set:

```{r}
# Calculate mean of steps taken per day
newMeanDailySteps <- format(round(mean(newDailySteps$steps), digits=2), nsmall=2)

# Calculate median of steps taken per day
newMedianDailySteps <- as.integer(median(newDailySteps$steps))
```

* The new mean for daily steps = `r newMeanDailySteps`. The previous value was `r meanDailySteps`. This is a difference of `r as.numeric(meanDailySteps) - as.numeric(newMeanDailySteps)`
* The new median for daily steps = `r newMedianDailySteps`. The previous value was `r medianDailySteps`. This is a difference of `r medianDailySteps - newMedianDailySteps`

Replacing the missing values with the mean for the same interval does not appear to significantly affect the results.

###Are there differences in activity patterns between weekdays and weekends?
* Figure out which days are weekends
```{r}
# Convert date field to POSIXlt data type
data$date <- as.POSIXlt(data$date)

# Add weekend variable
data$weekend <- ifelse(data$date$wday %in% c(0, 6), "Weekend", "Weekday")

# Summarize data by weekend vs weekday
sumData <- ddply(data, c("interval", "weekend"), summarize, meanSteps=mean(steps))
```

* Graph data
```{r}
library(lattice)
xyplot(data$steps ~ data$interval | data$weekend, layout = c(1,2), type="l", xlab="Interval", ylab="Steps")
```