#Peer Assessment 1
```{r echo=FALSE}
# Load libraries
library(plyr)
```

###Loading and preprocessing the data

* First we need to load the activity data that we will be analyzing
```{r}
# Read data using read.csv()
data <- read.csv("activity.csv", na.strings=c("NA"))
```

### What is the mean total number of steps taken per day?

* Calculate the total number of steps taken per day:
```{r}
# Remove NA values
data <- data[is.na(data$steps)==0,]

# Summarize steps by day using plyr library
dailySteps <- ddply(data, .(date), summarize, steps=sum(steps))
```

* Histogram of steps taken per day:

```{r echo=FALSE}
hist(dailySteps$steps)
```

* Calculate the mean and median of the total steps taken per day

```{r}
# Calculate mean of steps taken per day
mean(dailySteps$steps)

# Calculate median of steps taken per day
median(dailySteps$steps)
```

### What is the average daily activity pattern?

* Calculate the average number of steps taken per 5 minute interval
```{r}
# Summarize steps per interval
intervalSteps <- ddply(data, .(interval), summarize, steps=mean(steps))
```

* Plot of average number of steps taken per interval
```{r echo=FALSE}
plot(intervalSteps$interval, intervalSteps$steps, type="l")
```

```{r echo=FALSE}
maxInterval <- head(intervalSteps[order(-intervalSteps$steps),], 1)
```

* The 5 minute interval with the highest average number of steps is **`r maxInterval$interval`** with an average of **`r maxInterval$steps` steps**.


